Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-07-20 demystify.tex
services

          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
          \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
          \item Sometimes adcom members do dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
          \item Being a minority can help (\autoref{sec:urm}).
          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}). Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
        \end{itemize}

@@

In most cases we make decision independently and do not talk to each other.
However, we do discuss when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.  We also chat to each other for interesting or strong applications that we think should be admitted, e.g., how to recruit this student or who should be the adviser. If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.  In short, while we do not talk to each other for most applications, we do for some.

>>>>

          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
          \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
          \item Sometimes adcom members dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
          \item Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications (\autoref{sec:read-all}).

          \item Being a minority can significantly help your application (\autoref{sec:urm}).
          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}). Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
        \end{itemize}

@@

In most cases adcom members, even those reviewing the same application, make decision independently and do not talk to each other.
However, in some rare cases we do talk to each other.  For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.  We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser. 
If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.

\section{Do adcom members read all applications?}\label{sec:read-all}

We only review those that are assigned to us. Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).

For example, I only review those that are assigned to me, which is typically around 50--70 applications for the Fall admission. 
I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam). But usually, even if I look at them, I do not get involved in their evaluation.
